### 2020-10-19 supp.tex
Start editing supplementary material

\section{Supplementary material}
In this section, we discuss how to numerically compute the shadowing trajectories of the Lorenz map. One of the motivations for approximating the Lorenz map is that having a closed form expression is necessary in the algorithm we discuss for numerical shadowing solution computation. The program used to generate Figure 
\ref{fig:lorenz_map} is in