### 2020-10-19 supp.tex
Add grebogi citations

\section{Supplementary material}
\subsection{The Lorenz Map}
\subsection{Approximation of the Lorenz Map}

\subsubsection{Computing shadowing solution of the Lorenz map}
In this section, we discuss how to numerically compute the shadowing trajectories of the Lorenz map. One of the motivations for approximating the Lorenz map is that having a closed form expression is necessary in the algorithm we discuss for numerical shadowing solution computation. The program used to generate Figure 
\ref{fig:lorenz_params_scaled} is in \verb+lorenz_map/shadow.py+. The main logic behind the numerical procedure to compute the shadowing solution can be found in that script, within the function \verb+shadow+, which calls \verb+find2+.