### 2020-12-13 highlights.tex
Add highlights and edit intro. Final JCP version.

\documentclass[12pt]{article}
\title{Research highlights}
\begin{document}
\maketitle
\begin{itemize}
\item This paper highlights through rigorous examples that numerical solutions of chaotic dynamical systems are not guaranteed to produce the correct statistics or long-term behavior;
\item We show through examples that it is theoretically possible to significantly alter the statistics of a chaotic system by minor parameter perturbations.
\item The nonphysicality of shadowing orbits also implies that methods to compute parametric derivatives of statistics in chaotic systems, based on shadowing, may produce incorrect results.
\end{itemize}
\end{document}