### 2021-04-23 letter/summaryOfRevisions.tex
Add summary of revisions.

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1.1in]{geometry}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{verbatim}
\usepackage[textsize=tiny]{todonotes}
\usepackage{amsmath,amssymb}
\usepackage{fancyhdr}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\pagestyle{fancy}
\lhead{Response to the Reviewers}
\rhead{30th October 2020}
\lfoot{\href{mailto:nishac@mit.edu}{Nisha Chandramoorthy} and \href{mailto:qiqi@mit.edu}{Qiqi Wang}}
\cfoot{}
\rfoot{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\definecolor{azure}{rgb}{0.0, 0.5, 1.0}
\definecolor{burgundy}{rgb}{0.5, 0.0, 0.13}
\newcommand{\reviewerOne}[1]{{\color{burgundy}\textbf{#1}}}
\newcommand{\reviewerTwo}[1]{{\color{azure}\textbf{#1}}}

\title{}
\author{}
\date{23rd April 2021}
\setlength{\marginparwidth}{2.1cm}
\begin{document}

\begin{center}
		\Large{\textbf{Manuscript JCOMP-D-20-01958: Response to Reviewers}}
\end{center}
\medskip
We would first like to thank both the reviewers for investing their time to read our work carefully and provide us with thoughtful feedback. We have revised our work to incorporate all of both the reviewers' suggestions. Here is a summary of the changes we have made:
\begin{itemize}
    \item We have added citations to works pointed out to us by Reviewer 1 that strengthen the conclusions of this paper, especially on the implications of our results for the statistical inaccuracy of numerical solutions to chaotic systems. A brief explanation of the cited works is added to section 5, and is highlighted in \reviewerOne{red}.
    \item We have added a couple of paragraphs in section 3.2 (in \reviewerOne{blue}) to explain the sense in which we refer to periodic solutions as nonphysical. This is in response to a comment by Reviewer 1 on periodic solutions having physical interpretations.
    \item 
\end{itemize}
\end{document}