### 2021-05-06 highlights.tex
highlights modify and change supp to appendix

\documentclass[12pt]{article}
\title{Research highlights}
\begin{document}
\maketitle
\begin{itemize}
\item Shadowing orbits may not be typical orbits or physical solutions
\item This atypicality implies shadowing-based sensitivity can be biased.
\item Numerical solutions of chaotic dynamical systems can be statistically inconsistent;
\item Minor parameter changes can significantly alter the statistics of a chaotic system.

\end{itemize}
\end{document}