### 2021-05-06 highlights.tex
Changes made

\documentclass[12pt]{article}
\title{Research highlights}
\begin{document}
\maketitle
\begin{itemize}
\item Shadowing orbits may not be typical orbits or physical solutions
\item This atypicality implies shadowing-based sensitivity can be biased
\item Numerical solutions of chaotic dynamical systems can be statistically inconsistent
\item Minor parameter changes can significantly alter the statistics of a chaotic system

\end{itemize}
\end{document}